General constraints on magic distillation protocols from statistical mechanics

### 2021-06-04 magic_paper.tex
expanded argument

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $\alpha_1 / \alpha_2$, with even $\alpha_1 \geq 0$ and odd $\alpha_2 > 0$.
Note that for any quantum state $\rho$, $H_1$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}
\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]}

@@

\begin{theorem}
	Let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions such that $\w \succ \w'$.
	Then, $H_\alpha(\w) \leq H_\alpha(\w')$, where $\alpha$ is such that $\hspace{-1pt}H_\alpha(\x)\hspace{-2pt}$ is well-defined for all quasi-distributions $\x$.
\end{theorem}
\begin{proof}
Given a fixed $\alpha$ for which $H_\alpha(\x)$ is well-defined for all quasi-distributions $\x$, the function $H_\alpha$ is real-valued, smooth on the interior of $\mathcal{Q}^\downarrow$ and symmetric on permutation components of $\x$.

Consider the partial derivatives
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} = \frac{\alpha w_i^{\alpha-1}}{(1-\alpha)\sum_k{w_k^\alpha}} 
\end{equation}
Whenever $w_i \geq w_j$, we have that 
\begin{equation}
	\frac{w_i^{\alpha-1}}{1-\alpha} \leq \frac{w_j^{\alpha-1}}{1-\alpha},
\end{equation}
\nick{wrong} independently of whether $\alpha < 1$ or $\alpha > 1$, which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$.
\end{proof}

We can now derive the following entropic bounds on magic state transformations. \nick{to be extended (1) from unital processes to general relative majorization, (2) by dropping the condition $\alpha > 1$}

\begin{theorem}
	Consider a general magic state distillation process $\rho^{\otimes n} \rightarrow \rho'^{\otimes n'}$	in the unital fragment.
	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho') \right)},
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{theorem}
\begin{proof}
	We can regularise the number of output copies in the state distillation by tensoring in $(n-n')$ copies of the (free) maximally mixed state $\id/3$, so that we instead consider the process
\begin{equation}
	\rho^{\otimes n} \rightarrow \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')}.
\end{equation}
Since we have that $W_{\rho^{\otimes n}}(\bmx) \succ_{\id/3} W_{\rho'^{\otimes n'}}(\bmx)$, by Schur-concativity of $H_\alpha$ we obtain
\begin{equation}
	H_\alpha \left( \rho^{\otimes n} \right) \leq H_\alpha \left( \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')} \right).
\end{equation}
By additivity of $H_\alpha$, we can rewrite the inequality as
\begin{equation}
	n H_\alpha \left( \rho \right) \leq n' H_\alpha \left( \rho' \right) + (n - n') H_\alpha \left( \id/3 \right).
\end{equation}
Now, $H_\alpha \left( \id/3 \right)$ is the maximum value of $H_\alpha(\rho)$ for unital processes because every state $\rho$ majorizes $\id/3$, so we can finally re-arrange the inequality to
\begin{equation}
	\frac{n'}{n} \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho' \right)}.
\end{equation}
\end{proof}

@@

We have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{\alpha} \\
&= \frac{1}{(\alpha-1) \sum_i w_i^{\alpha}}\alpha w_k^{\alpha-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^\alpha$ for simplicity, that
\begin{align}
 \frac{1}{(\alpha-1)S}\alpha w_i^{\alpha-1}\ge \frac{1}{(\alpha-1)S}\alpha w_{i+1}^{\alpha-1},
\end{align}
for each $i=1,2,\dots, N-1$. We have that $S >0$, since $a$ is even, and assume that $\alpha >1$ so that
\begin{equation}
w_i^{\alpha-1}\ge w_{i+1}^{\alpha-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $\alpha = a/b > 1$ we have $\alpha-1 = (a-b)/b = a'/b >0$ where $a'$ is an even integer and $b$ an odd integer. We now assume that $\w \in \mathcal{Q}^\downarrow$ so that $ w_i \ge w_{i+1}$. \ddd{[Wait I'm not sure the inequalities work...]}
\end{proof}

>>>>

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $\alpha_1 / \alpha_2$, with even $\alpha_1 \geq 0$ and odd $\alpha_2 > 0$.
Note that for any quantum state $\rho$, $H_1$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_vs_eps_a.pdf}
    \caption{\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]} Here, the exponent is $\alpha = even/odd$ for $1.10 \leq \alpha \leq 1.50$. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(0)) = 0$.
    }
    \label{fig:lcs}
\end{figure}

@@

\begin{theorem}\label{thm:HSchur}
	Let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions such that $\w \succ \w'$.
	Then, $H_\alpha(\w) \leq H_\alpha(\w')$, where $\alpha = \frac{2a}{2b+1}$ for any positive integers $a,b$ with $a > b$.
\end{theorem}
\begin{proof}
We have $\alpha = \frac{2a}{2b+1} \geq \frac{2b+2}{2b+1} > 1$ and $\alpha$ is a rational with even numerator and odd denominator, so $\sum_i x_i^\alpha > 0$ and $H_\alpha(\x)$ is well-defined for all quasi-distributions $\x$.

Given a fixed $\alpha$ for which $H_\alpha(\x)$ is well-defined, the function $H_\alpha$ is real-valued, smooth on the interior of $\mathcal{Q}^\downarrow$ and symmetric on permutation components of $\x$.

Consider the partial derivatives
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} = \frac{\alpha}{(\alpha-1)\sum_k{w_k^\alpha}}(-w_i^{\alpha-1})
\end{equation}
Whenever $w_i \geq w_j$, for $\alpha$ in the given form, we always have that $\alpha - 1$ is positive and of the form odd$/$odd so \nick{expand argument}
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$.
\end{proof}
Note that if $\alpha < 1$ or is of the form odd$/$even, the argument of~\cref{thm:HSchur} does not hold.

We can now derive the following entropic bounds on magic state transformations.
\begin{theorem}
	Consider a general magic state distillation process $\rho^{\otimes n} \rightarrow \rho'^{\otimes n'}$	in the unital fragment.
	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho') \right)},
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{theorem}
\begin{proof}
	We can regularise the number of output copies in the state distillation by tensoring in $(n-n')$ copies of the (free) maximally mixed state $\id/3$, so that we instead consider the process
\begin{equation}
	\rho^{\otimes n} \rightarrow \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')}.
\end{equation}
Since we have that $W_{\rho^{\otimes n}}(\bmx) \succ_{\id/3} W_{\rho'^{\otimes n'}}(\bmx)$, by Schur-concativity of $H_\alpha$ we obtain
\begin{equation}
	H_\alpha \left( \rho^{\otimes n} \right) \leq H_\alpha \left( \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')} \right).
\end{equation}
By additivity of $H_\alpha$, we can rewrite the inequality as
\begin{equation}
	n H_\alpha \left( \rho \right) \leq n' H_\alpha \left( \rho' \right) + (n - n') H_\alpha \left( \id/3 \right).
\end{equation}
Now, $H_\alpha \left( \id/3 \right)$ is the maximum value of $H_\alpha(\rho)$ for unital processes because every state $\rho$ majorizes $\id/3$, so we can finally re-arrange the inequality to
\begin{equation}
	\frac{n'}{n} \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho' \right)}.
\end{equation}
\end{proof}

@@

We have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{\alpha} \\
&= \frac{1}{(\alpha-1) \sum_i w_i^{\alpha}}\alpha w_k^{\alpha-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^\alpha$ for simplicity, that
\begin{align}
 \frac{1}{(\alpha-1)S}\alpha w_i^{\alpha-1}\ge \frac{1}{(\alpha-1)S}\alpha w_{i+1}^{\alpha-1},
\end{align}
for each $i=1,2,\dots, N-1$. We have that $S >0$, since $a$ is even, and assume that $\alpha >1$ so that
\begin{equation}
w_i^{\alpha-1}\ge w_{i+1}^{\alpha-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $\alpha = a/b > 1$ we have $\alpha-1 = (a-b)/b = a'/b >0$ where $a'$ is an even integer and $b$ an odd integer. We now assume that $\w \in \mathcal{Q}^\downarrow$ so that $ w_i \ge w_{i+1}$. \ddd{[Wait I'm not sure the inequalities work...]} \nick{Look at proof of~\cref{thm:HSchur} and table below}

\begin{table}[h]
  \def\arraystretch{1.5}
  \centering
  \begin{tabular}{c|c|c|c||c|c}
    $w_i$ & $w_{i+1}$ & $\alpha$ & $\frac{w_i^{\alpha-1}}{\alpha-1} \geq \frac{w_{i+1}^{\alpha-1}}{\alpha-1}$ & $\alpha$ & $\frac{w_i^{\alpha-1}}{\alpha-1} \geq \frac{w_{i+1}^{\alpha-1}}{\alpha-1}$ \\[0.5ex]\hline
    $2$ & $1.5$  & $3$ & $2 > 1.125$ T & $1/3$ & $-0.94 > -1.14$ T \\
    $2$ & $-1.5$ & $3$ & $2 > 1.125$ T & $1/3$ & $-0.94 > -1.14$ T \\
    $2$ & $-2.5$ & $3$ & $2 < 3.125$ F & $1/3$ & $-0.94 < -0.81$ F \\[0.5ex]\hline
    $2/3$ & $1/3$  & $3$ & $0.22 > 0.06$ T & $1/3$ & $-1.97 > -3.12$ T \\
    $2/3$ & $-1/3$ & $3$ & $0.22 > 0.06$ T & $1/3$ & $-1.97 > -3.12$ T \\
    $2/3$ & $-4/3$ & $3$ & $0.22 < 0.89$ F & $1/3$ & $-1.97 < -1.24$ F \\[0.5ex]\hline
    $2$ & $1.5$  & $4$ & $2.67 > 1.125$ T & $2/3$ & $-2.38 > -2.62$ T \\
    $2$ & $-1.5$ & $4$ & $2.66 > -1.125$ T & $2/3$ & $-2.38 > 2.62$ F \\
    $2$ & $-2.5$ & $4$ & $2.66 > 5.21$ T & $2/3$ & $-2.38 > 2.21$ F \\
    $-2$ & $-2.5$ & $4$ & $-2.66 > -5.21$ T & $2/3$ & $2.38 > 2.21$ T \\[0.5ex]\hline
    $2/3$ & $1/3$  & $4$ & $0.10 > 0.01$ T & $2/3$ & $-3.43 > -4.33$ T \\
    $2/3$ & $-1/3$ & $4$ & $0.10 > -0.01$ T & $2/3$ & $-3.43 > 4.33$ F \\
    $2/3$ & $-4/3$ & $4$ & $0.10 > -0.79$ T & $2/3$ & $-3.43 > 2.73$ F \\
    $-1/3$ & $-2/3$ & $4$ & $-0.01 > -0.10$ T & $2/3$ & $4.33 > 3.43$ T \\
    $-1/3$ & $-4/3$ & $4$ & $-0.01 > -0.79$ T & $2/3$ & $4.33 > 2.73$ T
  \end{tabular}
  \caption{Exploring in what regimes the Schur-Ostrowski criterion holds.}
  \label{tab:test}
\end{table}
